%!TEX root = sketching-convex-ols.tex

The problem of handling the growing volume of data is of increasing
importance, particularly in terms of designing theoretically attractive
algorithms
which are efficient for the practitioner.
In light of this, matrix sketching has recently emerged as a method for
 approximating certain
linear algebra algorithms for popular data analysis primitives.
The main idea is that properties of a data matrix can be well-understood
through a
smaller summary of the data which is much easier to use in computations and
can offer huge computational savings.
Let $A \in \R^{n \times d}$ be the input sample-by-feature matrix with
associated target vector $b \in \R^n$ and assume that $n \gg d$.
Additionally, let $\mathcal{C}$ denote a set of convex constraints and the
task is to find

\begin{equation} \label{eq: convex-ols-problem}
  x_{OPT} = \argmin_{x \in \mathcal{C}} \frac{1}{2} \|Ax-b\|_2^2.
\end{equation}
\noindent In the large data setup we assume that $n$ is sufficiently large so
that solving Equation (\ref{eq: convex-ols-problem}) exactly is not possible.
As a result, one needs to exploit some notion of approximation to solve this
problem efficiently.
Two competing methods of approximation through matrix sketching are the
 \textit{sketch-and-solve} and \textit{iterative Hessian sketching} models which we
  detail below.



\begin{itemize}
  \item \textit{Sketch-and-solve}:
  The sketch-and-solve approaches output estimates found by solving

  \begin{equation}
    \hat{x} = \argmin_{x \in \mathcal{C}} \| S(Ax - b) \|^2
  \end{equation}
  by sampling a random linear transformation $S \in \R^{m \times n}$ from a
  sufficiently well-behaved distribution of matrices with $m \ll n$.
  Now that $m$ is much smaller than $n$, the $m \times d$ dimensional problem
  is small enough that it can be solved exactly.
  In fact, for appropriately chosen $S$, one can show that the \textit{
  cost} of a regression problem can be aproximated accurately in the sense
  that $\| A  \hat{x} - b \|^2 \le (1 + \eps) \|A x_{OPT} - b\|^2$ where
  $\eps$ is an accuracy parameter and the result holds with high constant
  probability.
  It is assumed that the projection dimension $m$ is chosen sufficiently small
  relative to $n$ that the smaller problem is efficiently solvable and that
  the main computational bottleneck is the time it takes to compute the
  summary $SA$.
  We detail the computational complexity of this approach in Section
  \ref{sec: preliminaries}.

  \item \textit{Iterative Hessian Sketching}:
  This approach exploits the quadratic program formulation of Equation
  (\ref{eq: convex-ols-problem}) and uses the random projections to accelerate
  expensive computations in the problem setup.
  In addition, one can argue that the summary $SA$ is acting as a preconditioner
  for the original problem.
  In contrast to the sketch-and-solve approach, the aim here is to define an
  iterative scheme through which one gradually refines the estimate in order
  to descend to the true solution of the problem outlined in Equation (\ref{eq: IHS}).

  \begin{equation} \label{eq: IHS}
    x^{t+1} = \argmin_{x \in \mathcal{C}}  \frac{1}{2} \|S^{t+1} A
    (x - x^t) \|^2 - \langle A^T (y - Ax^t), x - x^t \rangle
  \end{equation}
  The benefit of this approach is that rather than computing $\|Ax\|_2^2$ one
  instead computes $\| SAx \|_2^2$ which is sufficiently-well concentrated to enjoy
  convergence to the optimal solution through the iterative scheme.
  This is a huge computational saving when $SA$ has $m$ rows (which can be bounded
  independently of $n$) rather than the $n$ of $A$ to compute the quadratic form.

\end{itemize}

\textbf{Related Work.}
Within this family of convex constrained least squares problems are popular
data analysis tools such as ordinary least squares ($\mathcal{C} = \R^d$),
and penalised forms of regression: $\mathcal{C} = \{x : \|x\|_p \le t, p=1,2 \}$
among many others such as Elastic Net and SVM.
For problems such as unconstrained regression or LASSO, the time complexity of
solving the optimisation problem (i.e without the use of cross-validation to
choose hyperparameters) is  $O(nd^2)$.
There are various works which have studied these regression problems from both
the sketch-and-solve and IHS perspectives.



\noindent\textit{Sketch-and-solve}:
Provided that a sketch of a matrix is a \textit{subspace embedding} (defined
in Section \ref{sec: preliminaries}) which roughly requires that norms of vectors
are preserved under the transformation $S$, then the cost
of the regression problem can be approximated up to $(1 \pm \eps)$\textit{
relative error} \cite{woodruff2014sketching}.
Sarl\'{o}s used \textit{Fast Johnson Lindenstrauss Transforms}
to construct subspace embeddings to find
the first $o(nd^2)$ algorithm for solving unconstrained regression
 \cite{sarlos2006improved}.
 In addition, a similar idea was employed by Clarkson and Woodruff
\cite{clarkson2013low} to
generate a subspace embedding which could be found in time proportional to the
number of nonzeros in the matrix (denoted $\nnz{A}$ throughout).
Similar ideas have been extended to ridge regression in
\cite{avron2016sharper} as well as other forms of constrained regression again
found in \cite{woodruff2014sketching}.

A particular strength of the sketch-and-solve approach is that only one pass
of the data and
as such it is attractive for extremely large datasets.
Additionally, the \textit{cost}, or objective value, of a regression can be
approximated up to $(1 \pm \epsilon)$ relative error.
However, a draqback of the sketch-and-solve method is that although one can
accurately preserve the objective value of the original problem,
the projection onto a random subspace yields suboptimal estimators of the
solution vector to the original problem.
Therefore, despite these strengths, it is less well-known how the
 sketch-and-solve
approach fairs when new examples are added which is of considerable
importance to machine learning practitioners.
This has been addressed to some extent in \cite{price2017fast} where the
authors give an $\ell_{\infty}$ guarantee on the distance between the
approximate estimator and the optimal solution vector, yet also shown
was the existence of a pathological
regression instance for the CountSketch as opposed to the SRHT.
As a result, using the CountSketch within constrained regression problems
for solution recovery remains little used and given the potential speedup
from using this sketch, one which is possibly very useful.

A recent work of \cite{dahiya2018empirical} has explored the use of the
CountSketch method for a variety of problems.
Although the authors detail the efficacy of this sketch, they focus on
variants of the ordinary least squares problem, robust regression, and other
linear algebra algorithms in the sketch-and-solve model.
As such, the aim of our work is different as we focus solely on convex
constrained regression problems with a view to comparing the two competing
models as well as potential improvements (or deficiencies) of the CountSketch
compared to other random projections.


% Keep below for a two column paper
% \begin{align*}
%     x^{t+1} = \argmin_{x \in \mathcal{C}}  \frac{1}{2} \|S^{t+1} A
%     (x &- x^t) \|^2 \\
%       &- \langle A^T (y - Ax^t), x - x^t \rangle
% \end{align*}

\noindent\textit{IHS}:
Rather than simply solving $\min \|S(Ax-b)\|^2$ Pilanci and Wainwright
\cite{pilanci2016iterative}
proposed an iterative method based on solving the iterative
scheme defined in Equation (\ref{eq: IHS}).
The benefit of the IHS approach is that after sufficiently many iterates,
the output approximation, $\hat{x}$, has small error from the true
optimal solution.
This idea was later refined to sketch both large ($n$) and high dimensional
problems ($d$) through an iterative primal-dual approach in
 \cite{wang2017sketching} but applied specifically to the case of
  high-dimensional ridge regression:
 however, again, the faster CountSketch transform
was not studied.
In a similar line of work, the IHS method was extended to a wider class
of problems (i.e. those whose objective function is twice-differentiable and
convex) than the constrained least-squares regression
\cite{pilanci2017newton} and hence is beyond the scope of this paper.


\textbf{Contributions.}
Our work seeks to understand the benefits and limitations of the two
above competing sketching methods in convex constrained regression.
First we show theoretically that the CountSketch can be used in the iterative
framework developed by Pilanci and Wainwright.
Secondly, we show that the CountSketch also suffers from the same
 solution-approximation
suboptimality under the (standard) Gaussian design setting which motivates its
use in the iterative setup.
Empirically we demonstrate that the CountSketch performs comparably with
other sketching methods on various synthetic and real datasets:
importantly the required projection dimension behaves better than suggested by
the theory.
Then we show that, in line with the theory, CountSketch is practically a much
faster summary method as it can be computed while `streaming` the dataset.
We then give a series of experiments comparing the types of sketches within the
IHS approach and show that for this class of problems the CountSketch is a good
choice of random projection.
Although using a CountSketch as a preconditioner has been studied before
\cite{dahiya2018empirical}, its use
within the IHS framework is not understood and this work highlights that
significant
speedups can be found by exploiting the CountSketch transform.
Our work is timely as much of the current research on randomized sketching
 algorithms
has focused on the theory and there is still a gap between theoretical and
practical understanding which we aim to narrow.
